---
title: "lab1"
author: "Mahmoud Mousa Hamad"
date: "1/15/2020"
output: html_document
---

# Task 1

> Print working directory.

```{r}
getwd()
```

# Task 2

> Obtain first six lines of the data.

```{r}
ddt = read.csv("Lab1.csv")
head(ddt)

```

# Task 3

## Task 3.1

> What are the qualitative variables in “ddt”?

> ANSWER: RIVER and SPECIES

## Task 3.2

> What are the quantitative variables in “ddt”?

> ANSWER: MILE, LENGTH, WEIGHT, and DDT.

## Task 3.3

> How many SPECIES are in the ddt data set?

> ANSWER: 3

## Task 3.4

> Subset the ddt data set so that we see only the observations with SPECIES,  LMBASS and WEIGHT bigger than 800gms.

``` {r}
ddt[ddt$SPECIES %in% c("LMBASS") & ddt$WEIGHT > 800, ]
```

## Task 3.5

> Subset the ddt data set so that we see only those observations where the RIVER is SCM   and DDT bigger than 4.0.

``` {r}
ddt[ddt$RIVER %in% "SCM" & ddt$DDT > 4.0, ]
```

## Task 3: Clicker Questions Sub-Sections

### Q1

> Find the mean length of fish in the ddt data frame

```{r}
mean(ddt$LENGTH)
```


### Q2

> What is the standard deviation of the weight of fish in the ddt data frame?

```{r}
sd(ddt$WEIGHT)
```


### Q3

> Is this the plot of LENGTH vs WEIGHT from the ddt data frame?

``` {r}
plot(ddt$LENGTH, ddt$WEIGHT)
```

> ANSWER: YES

### Q4

> If v=1:20 what is the last value of v/20?

```{r}
v = 1:20
(v/20)[20]
```

# Task 4

## Task 4.1

> Make a table of the rivers.

```{r}
table(ddt$RIVER)
```


## Task 4.2

> Make a color coded barplot of the rivers.

```{r}
barplot(table(ddt$RIVER), col = c("red", "blue", "yellow", "green"))
```

## Task 4.3

> Make a table of the rivers crossed with the species of fish.

```{r}
my_table = table(ddt$RIVER, ddt$SPECIES)
my_table
```


## Task 4.4

> Make a barplot of the rivers crossed with the species of fish.

```{r}
barplot(my_table, legend = unique(ddt$RIVER), col = c("red", "blue", "green", "yellow"), args.legend = list(x = "topright"))
```

# Task 5

>	Make a pie chart of Fish SPECIES

```{r}
pie(table(ddt$SPECIES))
```


> Make a pie chart of RIVERS

```{r}
pie(table(ddt$RIVER))
```

# Task 6

> Make a boxplot of the variable DDT

```{r}
boxplot(ddt$DDT)
```

> Make a boxplot of the variable WEIGHT

```{r}
boxplot(ddt$WEIGHT)
```

> Make a boxplot of the variable LENGTH

```{r}
boxplot(ddt$LENGTH)
```

# Task 7

>	Make a coplot of LENGTH Vs WEIGHT given RIVER

```{r}
coplot(ddt$LENGTH ~ ddt$WEIGHT | ddt$RIVER)
```

>	Make a coplot of DDT Vs WEIGHT given SPECIES

```{r}
coplot(ddt$DDT ~ ddt$WEIGHT | ddt$SPECIES)
```

# Task 8

> Boxplot

```{r}
ggplot2::ggplot(ddt, ggplot2::aes(x = SPECIES, y = WEIGHT, z = RIVER, fill = RIVER)) +      ggplot2::geom_boxplot() +
ggplot2::ggtitle("Mahmoud Mousa Hamad")
```


> Violin plot

```{r}
library(ggplot2)
ggplot(ddt, aes(x = RIVER, y = LENGTH, z = SPECIES, fill = SPECIES)) + 
geom_violin()  +
ggtitle("Mahmoud Mousa Hamad")
```

> Scatter plot

```{r}
library(ggplot2)
ggplot(ddt, aes(x = WEIGHT, y = LENGTH, z = SPECIES, fill = SPECIES, color = SPECIES)) + 
geom_point()  +
ggtitle("Mahmoud Mousa Hamad")
```




